---
title: "DATA 405 Assignment 3"
author: "Rin Meng, 51940633"
output: pdf_document
---

## Question 1

(a) Quantile Function: The CDF of $V$ is:

$$
F_V(x) = 
\begin{cases}
1 - e^{-x^2}, & x > 0 \\
0, & x \leq 0
\end{cases}
$$

To find the quantile function, solve for $x$ in terms of $u$:

$$
F_V(x) = u \Rightarrow 1 - e^{-x^2} = u
$$ $$
e^{-x^2} = 1 - u
$$ $$
-x^2 = \ln(1 - u)
$$ $$
x = \sqrt{-\ln(1 - u)}
$$

Therefore, the quantile function $Q(u)$ for $V$ is:

$$
Q(u) = \sqrt{-\ln(1 - u)}
$$

Then, we can write the `rmyV` function using the quantile function:

```{r}
rmyV <- function(n) {
  u <- runif(n)
  x <- sqrt(-log(1 - u))
  return(x)
}
```

(b) The density function can be found by differentiating $F_v(x)$,

$$
    f_v(x) = \frac{d}{dx}F_V(x) = \frac{d}{dx}(1 - e^{-x^2}) = 2xe^{-x^2}, x > 0
$$

Then we can start coding for the simulation and plotting:

```{r}
dV <- function(x){
  ifelse(x > 0, 2 * x * exp(-x^2), 0)
}

set.seed(51940633)
simVals <- rmyV(10000)

hist(simVals, probability = TRUE, breaks = 50,
main = "Histogram of Simulated Values from V with Density Function Curve",
xlab = "Values", col = "lightblue", border = "black")

curve(dV(x), add = TRUE, col = "black", lwd = 2)
```

## Question 2

(a) The cumulative distribution of the function $X$ can be written as:

$$
F_X(x) = \int{f_X(x)} \, dx = \int{3x^2} \, dx = x^3
$$ So the CDF for $F_X(x)$ is:

$$
F_X(x) = 
\begin{cases}
x^3, & x \in [0, 1] \\
0, & x < 0 \\
1, & x > 1
\end{cases}
$$

(b) Now we can write the quantile function $Q(u)$, found by solving $F_X(x) = u$ for $x$, where $u$ is a uniform random variable on $[0, 1]$. $$
    F_X(x) = u \Rightarrow x^3 = u
    $$ $$
    x = u^{1/3}
    $$ So then the quantile function $Q(u)$ is, $$
    Q(u) = u^{1/3}
    $$ Now we can write the R function to generate random variates:

```{r}
rmyX <- function(n){
  u <- runif(n)
  x <- u^(1/3)
  return(x)
}
```

Now we define the PDF function for x, and simulate the distribution:

```{r}
dX <- function(x){
  ifelse(x >= 0 & x <= 1, 3 * x^2, 0)
}

set.seed(51940633)
simVals <- rmyX(10000)

hist(simVals, probability = TRUE, breaks = 50,
main = "Histogram of Simulated Values from X with PDF Curve",
xlab = "Values", col = "lightblue", border = "black")

curve(dX(x), add = TRUE, col = "black", lwd = 2)
```

## Question 3

(a) The CDF of $Y$ is given as:

$$
G(y) = pF_V(y) + (1 - p)F_X(y)
$$

Where $F_v(y) = 1 - e^{-y^2}$ is from question 1 and

$$
F_X(y) = 
\begin{cases}
y^3, & x \in [0, 1] \\
0, & y < 0 \\
1, & y > 1
\end{cases}
$$

from question 2.

(b) We can now generate random variates from Y,

```{r}
rmyY <- function(n, p) {
  # Step 1: Determine which distribution to sample from
  from_V <- rbinom(n, 1, p)  # Vector of 0s and 1s, where 1 means sample from V
  
  # Step 2: Generate samples from the appropriate distribution
  samples <- numeric(n)
  samples[from_V == 1] <- rmyV(sum(from_V == 1))  # Generate from V where from_V == 1
  samples[from_V == 0] <- rmyX(sum(from_V == 0))  # Generate from X where from_V == 0
  
  return(samples)
}
```

(c) Simulate the distribution of $Y$, for cases where p = 0.4:

```{r}
gY <- function(y, p) {
  p * (2 * y * exp(-y^2)) + (1 - p) * (ifelse(y >= 0 & y <= 1, 3 * y^2, 0))
}

set.seed(51940633)  # for reproducibility
simulated_values_Y <- rmyY(10000, p = 0.4)

hist(simulated_values_Y, probability = TRUE, breaks = 50,
     main = "Histogram of Simulated Values from Y with PDF Curve",
     xlab = "Values", col = "lightblue", border = "black")

curve(gY(x, p = 0.4), add = TRUE, col = "black", lwd = 2)

```
